%%% intro.tex
\section{Introduction}
\label{sec:intro}
%%%% 
Maximal repeats are string features widely used in bioinfomatics, which are defined as such substrings of a string that cannot be extended to the left or to the right without losing their occurrences. 
This paper studies the problem of enumerating all of $\mu$ distinct maximal repeats in a string of length $n$ using the suffix array (SA). 
Related to this problem, fundamental parameters of a string with length $n$ are the numbers $\mu$ and $e_R$ of maximal repeats and their right extensions, satisfying that $\mu \le e_R\le n$. 
It is well-known that these parameters $\mu$ and $e_R$ can be much smaller than the string length $n$, even logarithmically, for highly-repetitive strings. 

Since all of the previous algorithms for maximal repeats based on SA or BWT were known to require at least linear time in the text length $n$, it is a natural question whether it can be done in runing time in a way parameter sensitive to either $\mu$ or $e_R$, obtaining faster enumeration for highly-repetitive texts. 
As a main result, we show that the set of all maximal repeats in a string can be enumerated in $O(e_R)$ time and $O(\sigma^2 \log e_R)$ working space by a simple algorithm based on the suffix array with as auxiliary data structures of $O(n)$ space, namely, the inverse suffix array (ISA) and the range-minima structure on the longest common prefix (LCP) array. This result improves on the previous algorithms with $O(n)$ time complexity based on either the SA or BWT array of $O(n)$ space for highly-repetitive strings when $e_R = o(n)$. 

Furthermore, we also show that all maximal repeats can be enumerated in $O(e_R \;\textrm{polylog}(n))$ time and space on the top of existing compressed text indexing structures.
To the best of our knowledge, it is the first result to simultaneously achieve the time and space linear in $e_R$ for enumeration of all distinct maximal repeats with polylogarithmic factor in $n$. 

Technically, to obtain $O(e_R)$ running time, we devise a simple and modular algorithm by combining the top-down traversal of the suffix tree and constant time left-maxiamlity test in a novel way, which is not known before. To clearlify the situation around $O(e_R)$ computation of all maximal repeats of a string, we give a brief survey of existing algorithms for enumerating maximal repeats.

Overall, we obtained a simple and fastest algorithm for maximal repeats which is easy to implement and to be used for the real world sequence analysis, while it achieve the best simultaneous time and space complexities based on either the suffix arrays or the BWT arrays, comparable to one with the CDAWG, which is a fundamental compact data structure for maximal repeats. 

%%% EOF
